Exploratory Analysis

### 2023-11-30 Code/Draft1Rmd.Rmd
Updated limitation

#Run a synthetic control with the predictor variable being combined overdose deaths. The two states Washington and 
#California will serve as the donor pool states. Variables used will be smokers, percent fair or poor health, excessive drinking, and unemployed.
#The training data will be from 2016-2020, with 2021 and 2022 being the treatment outcomes section.
dataprep.out <-
  dataprep(health_ranking_yearly,
           predictors = c("smokers", "percent_fair_or_poor_health", "excessive_drinking"),
           dependent = "combined_overdose_deaths",
           unit.variable = "state_num",
           time.variable = "year",
           unit.names.variable = "X.1",
           treatment.identifier = 1,
           controls.identifier = c(2,3),
           time.predictors.prior = c(2016:2020),
           time.optimize.ssr = c(2016:2020),
           time.plot = 2016:2022,
  )

synth.out <- synth(dataprep.out)
print(synth.tables <- synth.tab(
  dataprep.res = dataprep.out,
  synth.res = synth.out)
)

@@

### Limitations 
It is a challenge to find a single cause for an impact on drug decriminalization and its relationship to drug overdose deaths. Since multiple factors can contribute to drug-related outcomes. There has been some evidence linked to the Covid-19 pandemic contributing to drug use, well-being, and mental health. There is limited generalizability to this study as findings from synthetic control analysis are specific to the context of Oregon drug policies and may not be generalized to other regions or policies. There is also limited data post treatment period (2 years). Additionally, the synthetic Oregon may not be a good fit. This will be further analyzed throughout the semester. As well as continuing to fully understand the statistical basis for synthetic control better.

>>>>

#Run a synthetic control with the predictor variable being combined overdose deaths. The two states Washington and 
#California will serve as the donor pool states. Variables used will be smokers, percent fair or poor health, excessive drinking, and unemployed.
#The training data will be from 2016-2020, with 2021 and 2022 being the treatment outcomes section.
dataprep.out <-
  dataprep(health_ranking_yearly,
           predictors = c("smokers", "percent_fair_or_poor_health", "excessive_drinking"),
           dependent = "combined_overdose_deaths",
           unit.variable = "state_num",
           time.variable = "year",
           unit.names.variable = "X.1",
           treatment.identifier = 1,
           controls.identifier = c(2,3),
           time.predictors.prior = c(2016:2020),
           time.optimize.ssr = c(2016:2020),
           time.plot = 2016:2022,
  )
synth.out <- synth(dataprep.out)
print(synth.tables <- synth.tab(
  dataprep.res = dataprep.out,
  synth.res = synth.out)
)

@@

### Limitations 
It is a challenge to find a single cause for an impact on drug decriminalization and its relationship to drug overdose deaths. Since multiple factors can contribute to drug-related outcomes. There has been some evidence linked to the Covid-19 pandemic contributing to drug use, well-being, and mental health. There is limited generalizability to this study as findings from synthetic control analysis are specific to the context of Oregon drug policies and may not be generalized to other regions or policies. There is also limited data post treatment period (2 years). Additionally, the synthetic Oregon may not be a good fit.
